---
title: "Untitled"
format: html
editor: visual
---

```{r}
library(fpp3)
```

# INSTRUCTIONS:

Complete the code snippets within this notebook with your code.

- DO NOT ADD ADDITIONAL CODE SNIPPETS. You can make the existing code snippets as long as you need, but you may not add new.
- DO NOT ALTER THE STRUCTURE OF THE NOTEBOOK.
- Stick to the syntax style used in class and use solely the library `fpp3`. This is the only library you need.

Groups that fail to adhere to these instructions will be penalized.

# Part 1

Decompose the time series below using classical decomposition. Do it using the function `classical_decomposition` from the fpp3 library as well as implementing the algorithm from scracth.

```{r}
a10 <- 
  # 1. Dataframe containing monthly medicare expenses per type of drug
  PBS %>%
  
  # 2. Filter for antidiabetic drigs
  filter(ATC2 == "A10") %>%
  
  # 3. select subset of columns
  select(Month, Concession, Type, Cost) %>%
  
  # 4. Add the total expenditure per month
  index_by(Month) %>% 
  summarise(TotalC = sum(Cost)) %>%
  
  # 5. Scale to millions
  mutate(Cost = TotalC / 1e6)

# 6. Plot
autoplot(a10, Cost) +
  labs(y = "$ (millions)",
       title = "Australian antidiabetic drug sales")
```

## 1.1

Using the function `classical_decomposition`, obtain the components. Store them in an object called class_dcmp1

```{r}
class_dcmp1 <- 
  # YOUR CODE GOES HERE
```

# 1.2

Obtain the same decomposition manually using the algorithm described in notebook 04_D. Store all the results in different columns of a dataframe (tsibble) called `manual_dcmp1`. Follow the steps below closely.

# 1.2.1 Obtain the trend using moving averages

```{r}
manual_dcmp1 <- 
  # Your code goes here
```

# 1.2.2 Detrend the time series

```{r}
manual_dcmp1 <- 
  # Your code goes here
```

# 1.2.3 Obtain the unadjusted seasonal component

Obtain the unadjusted seasonal component. Store it in a dataframe called `df_seasonal`, in a column called `s_unadj`

```{r}
# Any additional code you require goes here


df_seasonal <- 
  # Your code goes here
```

# 1.2.4

Adjust the seasonal component as indicated in the notebook 04_D_ClassicalDecomposition_fromScratch. Store it within `df_seasonal` in a new column called `seasonal_c`.

```{r}
# Any additional code to be added goes here:

# Update df_seasonal
df_seasonal <- 
  # Your code goes here
```

# 1.2.5

Import the seasonal component from `df_seasonal` into `manual_dcmp1` using a `left_join`

```{r}
manual_dcmp1 <- 
  left_join(... # TODO: complete this code)
```

# 1.2.6

Compute the remainder component

```{r}
manual_dcmp1 <- 
  # Your code goes here
```

# 1.2.7

Depict your manual decomposition using gg_plot, emulating the example given in notebook 04_D.

```{r}
# Your code goes here
```
