---
title: "TS_Ind_Assignment_2_CrossValidation"
toc: true
toc-location: left
toc-depth: 6
self-contained: true
format: html
editor: source
params:
  print_sol: true
---

```{r setup, include=FALSE}
library(fpp3)
```

## Exercise

For this exercise use the quarterly number of arrivals to Australia from New Zealand, 1981 Q1 – 2012 Q3, from data set `aus_arrivals`.

```{r}
# Filter data:
nzarrivals <- aus_arrivals %>% filter(Origin == "NZ")
nzarrivals
```

**a. Make a time-plot of your data. Ensure that the grid indicates at least the end of each seasonal period.**

**b. Create a training set that withholds the last two years of available data for test purposes.**

**c. Fit the following models to your training dataset and create 8-step ahead forecasts:**

**d. Which model performs better on average over all forecast horizon with this particular train-test split?**

**e. Perofrm cross-validation instead of using a single training-test set split.**

To perform cross-validation consider the following:

* The smallest training dataset shall contain at least 50 observations.
* The training datasets shall increase by one observation at a time.

1. Do you come to the same conclusions? Explain why or why not.
