---
title: "7_2_SES_Exercise_Solutions"
author: ""
date: ""
output:
  bookdown::html_document2:
    number_sections: no
    toc: true
    toc_depth: 6
    toc_float: true
    toc_collapsed: false
    self_contained: true
    lib_dir: libs
params:
  print_sol: FALSE 
  hidden_notes: FALSE
---

```{r, warning= FALSE, message=FALSE}
library(fpp3)
```

# SES Exercise

### 1. Select the data corresponding to Argentina from the dataset `global economy`. In particular, we are going to focus on its exports, so you may extract them as well. Produce a simple time plot of the exports

```{r}
arg_exports <- global_economy %>%
                  filter(Country == "Argentina") %>%
                  select(Year, Exports)

arg_exports
```

### 2. Fit a simple exponential smoothing method `ETS(A,N,N)` to the series. Plot the forecast along with the fitted values.

```{r}

```

```{r}

```
```{r}

```

#### * **Do you think the ETS forecast method is appropriate for this data? Why? Why not?**

### 3. How many parameters does the model have? Extract their values.

```{r}

```

### 4. Compute the standard deviation of the residuals as (see equation in html format of the notebook). See session 6_2 for more details.

\begin{equation}
  \hat{\sigma} = \sqrt{\frac{1}{T-K}\sum_{t=1}^T e_t^2}
\end{equation}

* $K$ is the number of parameters estimated in the forecasting method
* $T$ is the length of the residuals, after removing NAs

```{r}

```
### 5. Compute a 95% prediction interval for the first forecast using $\hat{y} \pm 1.96s$, where s is the standard deviation of the residuals computed in the previous point. See lesson 6_2 for further guidance.

```{r}

```

```{r}

```

### 6. Extract the prediction interval produced by R for the first forecast (see session 6_2). How does your prediction interval compare to that?

```{r}

```


```{r}

```
### 7. Write your own function to implement simple exponential smoothing. The function should take arguments:

* `y` (the time series)
* `alpha` (the smoothing parameter $\alpha$)
* `level` (the initial lvel $l_0$)

The function should return the value of the constant flat forecasts of SES beyond the training data.

```{r}
my_ses <- function(y, alpha, level) {

  # Returns the forecast (final level) of simple exponential smoothing
  
  # ARGUMENTS:
  # @y: a vector containing the y values of our time series
  # @alpha: parameter alpha of simple exponential smoothing
  # @level: initial level of simple exponential smoothing.
  
  # RETURN:
  # returns the final level (i.e. the flat forecast) of simple exponential smoothing
  
  yhat <- numeric(length(y) + 1) # Prepares the output of the for loop
  
  yhat[1] <- level
  
  for (i in 2:(length(yhat))) {
    yhat[i] <- ## your code goes here
  }
  
  return(last(yhat))
}
```
