---
title: "TS - Graphs Async1 recovery"
format: html
editor: source
params:
  print_sol: false
  print_sol_adv: false
  hidden_notes: false
toc: true
toc-location: left
toc-depth: 6
self-contained: true
---

```{r}
library(fpp3)
```

# Exercise 1

The aus_livestock data contains the monthly total number of pigs slaughtered in Victoria, Australia, from Jul 1972 to Dec 2018. 

1. Use filter() to extract pig slaughters in Victoria between 1990 and 1995. 

```{r, include=params$print_sol}
pigs_vict <- 
  aus_livestock %>% 
    filter(Animal == "Pigs") %>% 
    filter(State == "Victoria") %>% 
    filter(year(Month) >= 1990 & year(Month) <= 1995)
```

2. Use autoplot() and ACF() for this data. How do they differ from white noise?

```{r, include=params$print_sol}
pigs_vict %>% autoplot()
```


```{r, include=params$print_sol}
pigs_vict %>%
  ACF() %>% 
  autoplot()
```

# Exercise 2

The aus_arrivals data set comprises quarterly international arrivals to Australia from Japan, New Zealand, UK and the US.

1. Use filter to retain only the time series corresponding to Japan

```{r, include=params$print_sol}
japan_arrivals <- 
aus_arrivals %>% 
  filter(Origin == "Japan")
```

2. Create a timeplot of the data and comment on its pattern (comment max 60 words)

```{r, include=params$print_sol}
japan_arrivals %>% 
  autoplot()
```

3. Use `gg_season()` and `gg_subseries()` on the series and comment on the output of each of these. Max comment per graph of around 100+ words.

```{r, include=params$print_sol}
japan_arrivals %>% 
  gg_season()
```

```{r, include=params$print_sol}
japan_arrivals %>% 
  gg_subseries()
```

